---
title: "STA 100 Discussion 5"
author: "Gianni Spiga"
date: "2023-10-26"
output: html_document
---

## R Problems

### 1. Generate data sets of $n$ observations from normally distributed random variables using $n = 10, 100, 1000, 10000$.

Generate data sets of $n$ observations from normally distributed random variables using $n = 10, 100, 1000, 10000$.

```{r}
# By setting a seed, we are able to make sure Rstudio keeps our results consistent when we randomly generate numbers
set.seed(123)

# We can take advantage of the rnorm function here
# Default is mean =0, sd = 1, but you can always edit this to your needs
n10 = rnorm(10, mean = 0, sd = 1) 
n100 <- rnorm(100)
n1000 <- rnorm(1000)
n10000 <- rnorm(10000)
```



### 2. For each of the generated data sets in part (a)
 - (a) construct a histogram.
 - (b) create normal probability plots.

```{r}
# For n = 10
hist(n10)
{
qqnorm(n10)
qqline(n10)
}

# For n = 100
hist(n100)
{
qqnorm(n100)
qqline(n100)
}

# For n = 1000
hist(n1000)
{
qqnorm(n1000)
qqline(n1000)
}

# For n = 10000
hist(n10000)
{
qqnorm(n10000)
qqline(n10000)
}
```
\\

We can see as our sample size $n$ increases, our distributions become much more normal in appearence! 

### 3. Load the data given in the file **student.csv**. Construct a histogram and a normal probability plot of (a) students’ height and (b) students’ pulse. Would you agree with the statement that the data was generated from a normally distributed random variable? Why or why not?

```{r}
# Make sure your data is placed in the right folder! 
student <- read.csv("student.csv")

# Head() function helps us print the first five rows of the data, for preview
head(student)

# Great way to get a quick understanding of the data
summary(student)
```

```{r}
# (a)
# Curly brackets help if you get an error trying to run one line at a time 
{
qqnorm(student$height, main = "(a) Normal Probability Plot for Student's Height")
qqline(student$height)
}

hist(student$height, col = "orange")
```


Most of the points match up along a straight line. It suggests students heights are approximately normally distributed.

```{r}
# (b)
{
qqnorm(student$pulse, main = "(b) Normal Probability Plot for Student's Pulse")
qqline(student$pulse)
}

hist(student$pulse, col = "green")
```

The points forms a (convex) curve instead of a straight line. It indicates a right-skewed distribution. It does not suggest student’s pulses are approximately not normally distributed.

Confused on how to interpret a QQplot? Here is a great Stack Exchange page [here](https://stats.stackexchange.com/questions/101274/how-to-interpret-a-qq-plot). 

## Bonus - Changing your RStudio Theme

You can customize your RStudio color layout with built-in RStudio themes! Here is how to do it. 

- On the top left, click on **Tools**

- Scroll down to **Global Options** 

- On the left, click on the **Appearence** option

- Scroll through the list of **Editor Themes** and click **Apply** once you have chosen one. Once you do, your RStudio will have a new look! If you would like to get into the spirit of Halloween, I recommend the "Merbivore" theme. 


