---
title: "Discussion 3 STA 100"
author: "Gianni Spiga"
date: "2023-10-12"
output: html_document
---

# R Portion 

On Canvas, you will find the file “GSK.csv”. The csv file has the following columns:

Column 1. sysbp: The systolic blood pressure of the subject (mmHg). 

Column 2. gender: The gender, with levels F and M. 

Column 3. married: Y if the subject was married, N if not.

Column 4. exercise: With levels L = low, M = medium, H = high.

Column 5. age: The age of the subject in years.

Column 6. stress: With levels LS = low, MS = medium, HS = high.

Column 7. educatn: With levels LE = low, ME = medium, HE = high.

```{r}
# Let's load that data in 
GSK <- read.csv("GSK.csv") # Make sure you put correct path if not in folder
GSK
```

## a.) 

```{r}
hist(GSK$sbp, main = "SBP for All Subjects")

```
## b.)

```{r}
boxplot(sbp ~ exercise, data = GSK, main = "SBP by Exercise Level")

```

## c.)

```{r}
stress.table <- table(GSK$stress)
stress.table

barplot(stress.table)

```

## d.) 
```{r}
SE.table <- table(GSK$exercise, GSK$stress)
SE.table

mosaicplot(SE.table)


```